sWGS validation/protocol tests

### 2019-06-11 Validation Test Report.Rmd
Updating validation processing

dirs = list(
  training = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient',
  val_1 = '~/Data/BarrettsProgressionRisk/Analysis/val_test/',
  protocol = '~/Data/BarrettsProgressionRisk/Analysis/protocol_test/',
  kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/',
  ci_protocol = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol/'
)

qc.cutoff = 0.011
```

@@

```{r}
get.png<-function(n,dir) {
  list.files(path=dir, pattern=n, full.names=T, recursive=T)
}

@@

@@

We later developed a computational cutoff. A cutoff was selected for the median variance across segments in a sample of 0.015.  Any sample above that is considered to have failed QC.  In the initial training set that results in `r nrow(failed)` samples failing QC.  A manual check confirms that these samples are overdispersed (especially the two from AH0254).  This is `r round(nrow(failed)/nrow(train)*100,2)`% of our samples.

`r kable(train %>% dplyr::summarise_at(vars(varMAD_median), funs(median,mean,sd,min,max)), caption='Summary statistics for the variance across all training samples') %>% kable_styling(bootstrap_options=c('bordered',full_width=F))`

@@

`r sval %>% kable(caption='Summary of variance per batch. Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > cutoff), color='red')`

@@

@@

```{r}
ggplot(melt(pass.fail, id.vars = 'group'), aes(group, value,group=group,fill=variable)) + geom_col(position = position_stack()) + labs(y='Ratio of samples', x='') + theme_bw()
```

>>>>

dirs = list(
  #training = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient',
  val_1 = '~/Data/BarrettsProgressionRisk/Analysis/val_test/',
  protocol = '~/Data/BarrettsProgressionRisk/Analysis/protocol_test/',
  kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/',
  ci_protocol = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol/',
  old_vs_new = '~/Data/BarrettsProgressionRisk/Analysis/sujath_oac/',
  old_progressors = '~/Data/BarrettsProgressionRisk/Analysis/old_progressors/'
)

get.png<-function(n,dir) {
  list.files(path=dir, pattern=n, full.names=T, recursive=T)
}

qc.cutoff = 0.011
pass.fail = tibble()
```

@@

```{r, eval=F}

@@

<!-- We later developed a computational cutoff. A cutoff was selected for the median variance across segments in a sample of 0.015.  Any sample above that is considered to have failed QC.  In the initial training set that results in `r nrow(failed)` samples failing QC.  A manual check confirms that these samples are overdispersed (especially the two from AH0254).  This is `r round(nrow(failed)/nrow(train)*100,2)`% of our samples.

`r kable(train %>% dplyr::summarise_at(vars(varMAD_median), funs(median,mean,sd,min,max)), caption='Summary statistics for the variance across all training samples') %>% kable_styling(bootstrap_options=c('bordered',full_width=F))` -->

@@

@@

`r sval %>% kable(caption='Summary of variance per batch. Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > qc.cutoff), color='red')`

@@

## Sujath - Old vs New (ish) sample tests

Here we thought that we could see if the issue was older vs newer samples and we could compare that to the same patient with a frozen. These are all tumour samples (OAC) from OCCAMS.

```{r}
ci = do.call(bind_rows, lapply(list.files(dirs$old_vs_new, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types='cdddddl')
})) %>% select(sample, varMAD_median) %>% dplyr::mutate( Pass = varMAD_median<=qc.cutoff)

info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/sujath_oac/ffpe_frozen_samples.xlsx') %>%  
  dplyr::select(-matches('Adrienne|Notes')) %>% rename(Samplename='sample')

pass.fail = bind_rows(pass.fail, 
  ci %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(ci), Fail=length(which(varMAD_median>qc.cutoff))/nrow(ci), group='Old.v.New'))

plot.dir= '~/Data/BarrettsProgressionRisk/QDNAseq/sujath_oac/SLX-16753_16755_16758/'
proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/sujath_oac/'

ci.tbl = ci %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
 dplyr::mutate(
   Sample = text_spec(sample, link=sample),
#   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
   `Computational QC` = cell_spec(Pass, "html", background = ifelse(Pass == T, "green", "red"), color='white'),
   `Index Sequence` = text_spec(`Index Sequence`, tooltip=`sample`),   
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
  dplyr::select(`Patient ID`,`Index Sequence`, `Sample Type`, `Computational QC`, varMAD_median) %>% arrange(`Sample Type`)

rowsE = grep('FFPE',ci.tbl$`Sample Type`)
rowsC = grep('FFPE MRC',ci.tbl$`Sample Type`)
rowsN = grep('Frozen',ci.tbl$`Sample Type`)

ci.tbl %>%
  kable(format = "html", escape = F) %>% kable_styling("bordered", full_width = F) %>%
  group_rows("Cohort: FFPE", rowsE[1], rowsE[length(rowsE)],label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("Cohort: FFPE cut by AB", rowsC[1], rowsC[length(rowsC)],label_row_css = "background-color: #666; color: #fff;")  %>%
  group_rows("Cohort: Frozen", rowsN[1], rowsN[length(rowsN)],label_row_css = "background-color: #666; color: #fff;")

## Old progressors

Sujath and Adrienne extracted the progressor scrolls that had been cut about 12 months ago and left in tubes. These are likely to have oxidation issues making them potentially poor samples for inclusion, but it was worth checking. We included a number from Sujath's successful OAC extractions as controls.

```{r}
ci = do.call(bind_rows, lapply(list.files(dirs$old_progressors, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types='cdddddl')
})) %>% select(sample, varMAD_median) %>% dplyr::mutate( Pass = varMAD_median<=qc.cutoff)

info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/old_progressors/prog_val.xlsx') %>%  
  dplyr::mutate( sample = paste0(`SLX-ID`, '.', `Index Sequence`) ) %>% select(-matches('DNA')) %>% filter(!is.na(`SLX-ID`)) %>% dplyr::rename('Number' = 'Extraction Sample #')

pass.fail = bind_rows(pass.fail, 
  ci %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(ci), Fail=length(which(varMAD_median>qc.cutoff))/nrow(ci), group='Old Progressors'))

plot.dir= '~/Data/BarrettsProgressionRisk/QDNAseq/old_progressors/SLX-17965_17966/'
proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/old_progressors'

ci.tbl = ci %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
 dplyr::mutate(
   Sample = text_spec(sample, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),
   #`Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
   `Computational QC` = cell_spec(Pass, "html", background = ifelse(Pass == T, "green", "red"), color='white'),
   `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),   
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
  dplyr::select(`Patient ID`,`Index Sequence`, `Computational QC`, varMAD_median)

#rowsE = grep('Ellie',ci.tbl$Cohort)
#rowsC = grep('Cass',ci.tbl$Cohort)
#rowsN = grep('New', ci.tbl$Cohort)

ci.tbl %>%
  kable(format = "html", escape = F) %>% kable_styling("bordered", full_width = F)

@@

```{r, fig.height=6, fig.width=6}
cols = RColorBrewer::brewer.pal(3,'Set1')[c(1,3)]
ggplot(melt(pass.fail, id.vars = 'group'), aes(group, value,group=group,fill=variable)) + geom_col(position = position_stack()) + labs(y='Ratio of samples', x='') + theme_bw() + scale_fill_manual(values=c('Pass'=cols[2], 'Fail'=cols[1]), name='') 
```
